Adaptive Distance Metric Selection for Supervised Classification

### 2018-05-20 MS/thesis/tex 10 26/Adaptive_distance_metric_notes.tex
add cv

\title{Adaptive Distance Metric Selection for Supervised Classification}
\author{Krishnan Kumaran and Dimitri J. Papageorgiou \\
{\small Corporate Strategic Research}\\
{\small ExxonMobil Research and Engineering Company}\\
{\small 1545 Route 22 East, Annandale, NJ 08801 USA}\\
{\small dimitri.j.papageorgiou@exxonmobil.com} \\
}

@@

\section{Plot of number of outliers as function $\epsilon$}

\todo[inline]{this should be done before next meeting on 2016-10-27}

\begin{itemize}
\item 
  make a plots a number of outliers as function of epsilon
+ show also the datasets which you have used and so on
\item 
  try different datasets, also generate the two gaussian mixtures
\item   find out how to export figure to EPS from AIMMS
\item  change a copy of {\bf output scatter plot} to use the learnt distance measure
\end{itemize}

@@

\todo[inline]{Modify to incorporate assumption that $0 \leq d_{ij} \leq 1$ for all $i$ and $j$.}

\subsection{Similar to previous but we will have just one lambda}
\label{sec:mimax_oneLamnda}
\todo[inline]{FInish this soon and compare with previous approach}

@@

\newpage
\section{Nomenclature}
\begin{center}
\begin{tabular}{cl}
Decision variable & Definition \\
\hline
$d_{ij}$ & distance between points $i$ and $j$ \\
$w_{ij}$ & $d_{ij}y_{ij}$  \\
$y_{ij}$ & 1 if point $j$ is the closest neighbor to $i$; 0 otherwise  \\
$z_{i}$  & 1 if point $i$ is an outlier; 0 otherwise  \\
$\lambda_i$ & separation (difference) between closest non-neighbor and closest neighbor \\$\lambda_i$ & separation (difference) between closest non-neighbor and closest neighbor \\
$\v{B}$ & $\frac{p(p-1)}{2}$ \\ 
\end{tabular}
\end{center}

@@

\newpage 
\section{Progress Reports 2016-10-21}

\begin{itemize}
\item Loading instance works

\item Solving  \eqref{model:epsilon_lp_1} works now

\item Model \eqref{model:milp_with_outlier_detection}
works. I did it in the aimms code "MILP SECTION4" Part.

\end{itemize}
{\bf problems and questions:}
\begin{itemize}
\item queston 1 How to show the certain points in the plots?
\item question 2

\end{itemize}

{\bf Tasks to do }

\begin{itemize}
\item 1. Create new dataset - e.g.  - Guassian with diffrent mu and Sigma
\item 2. ammend the code to add new point in a.g. Scatter Point
\item 3. Implement Optimization of Distance Metrix                 ACCOMPLISHED
\item 4. Add the robust version for section 4. (have just one lambda) both Latex and AIMS (PROCESSING)
\item 5. Create a performance profiles -- example 100 datasets and comapre 3-4 distance measures 
\item 6. add new codes for replacing "a delta"  - to  "a |delta| "           ACCOMPLISHED
\item 7. add button to load instance          ACCOMPLISHED
\item 8. For Infesible problem we solved during the meeting why it was not infeasible                  ACCOMPLISHED

\end{itemize}

\newpage
\section{Progress Reports 2016-11-04}

\begin{itemize}
\item 1. Create new dataset - e.g.  - Guassian with diffrent mu and Sigma (ACCOMPLISHED)

\item 2. Add the robust version for section 4. (have just one lambda) in AIMMS (ACCOMPLISHED)

\item 3. Find out how to export figure to EPS from AIMMS (PROCESSING)

\item 4. Change a copy of output scatter plot to use the learnt distance measure (ACCOMPLISHED)

\end{itemize}
{\bf problems and questions:}
\begin{itemize}
\item 1. What should I do next? Higher dimension or something else?

\item 2. Still dont know how to draw a circle on a node in AIMMS.

\item 3. I covered the last file u updated by mistake, so I cant see the new theory. Would u please send it to me again? I will merge all of the files in one tex, thx.

\end{itemize}

\newpage
\section{Progress Reports 2016-11-11}

\subsection{Shatter outliers and see $\epsilon$ and Distance Matrices in LP approach}

First run the model in \eqref{model:epsilon_lp_1} in Aimms. 
We do these in Annulus Example 10.1 with 150 points, 2 classes, Probability of outliers = 0.05 and Noise = 0.5.

\begin{figure}[htbp]
\begin{center}
\includegraphics[scale=0.7]{Example_10_1_Origin.PNG}
%\vspace{-0.25in}
\caption{Example 10.1}
\label{fig:Example_10_1_Origin}
\end{center}
\end{figure}

After running the model in \eqref{model:epsilon_lp_1} in Aimms, we get a maximized $\epsilon$ =  0.001113532065 with outliers.

\begin{figure}[htbp]
\begin{center}
\includegraphics[scale=0.7]{Example_10_1_outliers_1.PNG}
%\vspace{-0.25in}
\caption{Example 10.1 Outliers}
\label{fig:Example_10_1_outliers_1}
\end{center}
\end{figure}

\begin{figure}[htbp]
\begin{center}
\includegraphics[scale=0.7]{Example_10_1_DistanceMatrix_1.PNG}
%\vspace{-0.25in}
\caption{Example 10.1 DistanceMatrix}
\label{fig:Example_10_1_DistanceMatrix_1}
\end{center}
\end{figure}

\newpage
Now we delete the outliers and do\eqref{model:epsilon_lp_1} in Aimms again. we get a maximized $\epsilon$ =  0.01597093404 with outliers again. This $\epsilon$ is much larger than $\epsilon$ =  0.001113532065 in Example 10.1.

\begin{figure}[htbp]
\begin{center}
\includegraphics[scale=0.7]{Example_10_2_outliers_1.PNG}
%\vspace{-0.25in}
\caption{Example 10.2 Outliers}
\label{fig:Example_10_2_outliers_1}
\end{center}
\end{figure}

\begin{figure}[htbp]
\begin{center}
\includegraphics[scale=0.7]{Example_10_2_DistanceMatrix_1.PNG}
%\vspace{-0.25in}
\caption{Example 10.2 DistanceMatrix}
\label{fig:Example_10_2_DistanceMatrix_1}
\end{center}
\end{figure}

\newpage
If we keep running the iterations, we get new outliers with new $\epsilon$ s and different Distance Matrix. In some cases, there will be always outliers after each iteration. Fortunately, In this case, we have only 3 iterations to get no outliers. In iteration 3, $\epsilon$ = 0.01597460757.

\begin{figure}[htbp]
\begin{center}
\includegraphics[scale=0.7]{Example_10_3_outliers_1.PNG}
%\vspace{-0.25in}
\caption{Example 10.3 Outliers}
\label{fig:Example_10_3_outliers_1}
\end{center}
\end{figure}

\begin{figure}[htbp]
\begin{center}
\includegraphics[scale=0.7]{Example_10_3_DistanceMatrix_1.PNG}
%\vspace{-0.25in}
\caption{Example 10.3 DistanceMatrix}
\label{fig:Example_10_3_DistanceMatrix_1}
\end{center}
\end{figure}

\newpage
\subsection{Put back outliers to new model}
We put back outliers to new model to see if they are still outliers.

\begin{figure}[htbp]
\begin{center}
\includegraphics[scale=0.7]{Example_10_3_OutliersBack_1.PNG}
%\vspace{-0.25in}
\caption{Example 10.3 Outliers Back}
\label{fig:Example_10_3_OutliersBack_1}
\end{center}
\end{figure}

Compaired to outliers in previous iterations, the previous outliers are still outliers with new Distance Matrices in iteration = 3.

{\bf problems and questions:}
\begin{itemize}
\item 1. If a case is "unfortunate" one(keeps popping out outliers in each iteration), how do we choose the value of $\epsilon$ to get a solution?

\item 2. I just show some case that "outliers are still outliers",but how do we guarantee others follow the same rule ?

\end{itemize}
\newpage

\section{Progress Reports 2016-11-18}

\begin{itemize}
\item 1. Fix the BUG of mistakes in showing the outliers (ACCOMPLISHED)

\item 2. Fix the mistake of formula using in AIMMS (ACCOMPLISHED)

\item 3. Do Iterations to refill the NonOutliers into the original dataset (Processing)

\end{itemize}
{\bf problems and questions:}
\begin{itemize}
\item 1. AIMMS CODE: How to clone a brand new set from a sub set of "a Set" without any effect when "a Set" changes? Stuck here.

\end{itemize}

@@

\section{Nomenclature}

>>>>

\title{Adaptive Distance Metric Selection for Supervised Classification}

@@

\todo[inline]{
My questions:

Value of epsilon matters only when it is nearly reaching “max epsilon”, so why do we care about it after this algorithm? I mean it will not change the final solution if we keep epsilon less equal than “max epsilon”/2.
}

@@

@@

\newpage
\section{Iterative LP Algorithm with Row and Column Generation}

A potential hurdle in the LP above is the huge number of constraints. For example, an instance with 1000
points, 10 classes, and 100 points per class would require 900,000 constraints for each ($i$,$k$) pair, while an
instance with 10,000 points, 10 classes, and 1000 points per class would require 9,000,000 constraints for
each ($i$,$k$) pair. Meanwhile, the fundamental theorem of linear programming asserts that if an optimal
solution exists, there exists an optimal basic feasible solution that will involve $n$ active constraints, where
$n$ = $N$ + ($p$ + 1)$p$/2 and $m$ = $N^2$ ($O$($N^2$)). Note that $n$ << $m$. There is likely to be a large amount of
degeneracy.

Our goal is to significantly reduce the number of constraints that must be explicitly included in the LP.
Let $N_A \subseteq   N$ be the set of active points considered in the LP. Let $S_i \subseteq  C_i$ and   $\overline S_i \subseteq \overline C_i $ for each point i 2 N.
Here, $S_i ( \overline S_i)$ represents the set of co-class (non-class) neighbors that are currently considered in the LP.

Let W be a positive scalar. Consider the following reduced LP that maximizes both $\epsilon$  and the minimum
distance variables $t_i$:

\begin{subequations} \label{model:big_lp_algorithm_1}
\begin{alignat}{4}
\max_{\v{a},\v{B},\v{d},\v{t}}~~& \sum_{i \in \mc{N}_A} t_i &&  \\
\st~~& t_i + \epsilon \leq d_{ik} && \qquad \forall i \in \mc{N}_A, k \in \overline{\mc{S}}_i \label{eq:big_lp_separation_constr_al_1} \\
    & t_i \leq d_{ij} && \qquad \forall i \in \mc{N}_A, j \in \mc{S}_i \label{eq:big_lp_dual_constr_al_1} \\
    & t_i \geq 0 & & \qquad \forall i \in \mc{N}_A \\
       & d_{ij} = \sum_{k=1}^p \delta_{ijk} a_{k} + \sum_{k=1}^p \delta_{ijk}^2 b_{kk} + 2 \sum_{k=1}^{p-1} \sum_{\ell=k+1}^p  \delta_{ijk} \delta_{ij\ell} b_{k\ell} && \qquad \forall i,j \in \mc{N}  \\
    & 0 \leq d_{ij} \leq 1 && \qquad \forall i,j \in \mc{N} \\
    & \v{a} \in \Re^p && \\ 
    & b_{k\ell} \in \Re && \qquad \forall k=1,\dots,p, \ell=k,\dots,p
\end{alignat}
\end{subequations}

\begin{figure}[htbp]
\begin{center}
\includegraphics[scale=0.15]{Fast_LP.png}
%\vspace{-0.25in}
\caption{LP Approach via Row and Column Generation (Fast LP)}
\label{fig:Fast_LP}
\end{center}
\end{figure}

\begin{algorithm}
%{\small
\caption{LP Approach via Row and Column Generation (Fast LP)}
\label{algo:Fast_LP}
\begin{algorithmic}[1]
\STATE Initialize: $\mc{N}_A= \mc{N}; O=\varnothing; S_i=\lbrace argmin\lbrace d_{ij}:j \in \mc{C}_i\rbrace \rbrace$; $\overline{S}_i=\lbrace argmin\lbrace d_{ik}:k \in \mc{N} \setminus \mc{C}_i\rbrace \rbrace$; Fix $\epsilon$ ; \textbf{B}= Euclidean distance metric;
\STATE DONE = False
\STATE (Re-)Solve model \eqref{model:big_lp_algorithm_1} for $B^*,d^*_{ij},t^*_i$. Fix the distance metric to $d^*_{ij}$.
\FOR{ $i \in \mc{N}_A$ }
	\STATE Set $\mc{V}_i= \lbrace j \in \mc{C}_i\setminus \mc{S}_i : t^*_i> d^*_{ij}\rbrace $. 
	\STATE Set $\overline{\mc{V}}_i= \lbrace k \in \overline{\mc{C}}_i\setminus \overline{\mc{S}}_i : t^*_i + \epsilon > d^*_{ik}\rbrace $.
 	\STATE Set $\mc{V}_i= \lbrace j \in \mc{C}_i\setminus \mc{S}_i : t^*_i> d^*_{ij}\rbrace $. 
\ENDFOR
\STATE ...
\STATE ...\todo[inline]{(need time to edit, I put a pic tells the algorithm down there)}

	\STATE Store the outliers and distance metric
\end{algorithmic}
%}
\end{algorithm}

\begin{figure}[htbp]
\begin{center}
\includegraphics[scale=0.8]{Fast_LP_fig.png}
%\vspace{-0.25in}
\caption{LP Approach via Row and Column Generation (Fast LP)}
\label{fig:Fast_LP_fig}
\end{center}
\end{figure}

\todo[inline]{
My questions:

1. Why NN, not 2-NN, 3-NN?  OK, it reduces the calculating time.

2. What if 3-NN or 6-NN performs better in accuracy? Do we need experiments on this?

3. Value of epsilon matters only when it is nearly reaching “max epsilon”, so why do we care about it ? I mean it will not change the solution if we keep epsilon less equal than “max epsilon”/2.
}

\newpage
\section{Outlier re-insertion}

\subsection{Outlier re-insertion in MILP}

.........\todo[inline]{(2 pages, Later I will put it on)}

\subsection{Outlier re-insertion with ranking of $R_i$}

Though the MILP problem is improved, when facing large scale of dataset, it will be out of the capability to solve in a short time. Here represents a new method with
$R_i =  {\min_{j \in \mc{C}_i} d_{ij}}/ {\min_{k \in \overline{\mc{C}}_i} d_{ik}}$

\begin{figure}[htbp]
\begin{center}
\includegraphics[scale=0.8]{Figure_R_i.png}
%\vspace{-0.25in}
\caption{Algorithm performing method with $R_i$}
\label{fig:Figure_R_i}
\end{center}
\end{figure}

Note that the iteration need to loop over several times to achieve the end. When the dataset is very large, it will be time consuming. To discard unnecessary procedures, introduce an improved re-insertion method:

\begin{algorithm}
%{\small
\caption{Improved Method of Outlier Re-insertion with Ranking of $R'_i$}
\label{algo:Improved_Method_of_Outlier_Re-insertion}
\begin{algorithmic}[1]
\STATE Compute $R_i$ for each outlier $i$; 
\STATE Identify all non-outliers $j$ for which $i$ is now the nearest co-class/non-class neighbor, i.e. all $j$ whose $R_i$ values are affected by insertion of $i$;
\STATE Define $R'_i= min (R_i  ,\min_j{R_j })$;
\STATE Sort outliers in the order of decreasing $R'_i$;
\STATE Insert the outliers according to decreasing ranking of $R'_i$;
\STATE Repeat steps 1-5 until no further re-insertion is possible, i.e. remaining outliers all have $R'_i<1$;

\end{algorithmic}
%}
\end{algorithm}

\todo[inline]{My question:

4.If Metric does not change after we put points back, so why do we do that? It’s meaningless.

5. About what to do next?

 (1) See how initialization of Matrix B influence the solving time in Fast LP.

 (2) Comparison of our method and others( k-nn ,svm…) with different dataset.

(3) Improving of the method (hardest)
}

\section{Nomenclature}
\begin{center}
\begin{tabular}{cl}
Decision variable & Definition \\
\hline
$d_{ij}$ & distance between points $i$ and $j$ \\
$w_{ij}$ & $d_{ij}y_{ij}$  \\
$y_{ij}$ & 1 if point $j$ is the closest neighbor to $i$; 0 otherwise  \\
$z_{i}$  & 1 if point $i$ is an outlier; 0 otherwise  \\
$\lambda_i$ & separation (difference) between closest non-neighbor and closest neighbor \\$\lambda_i$ & separation (difference) between closest non-neighbor and closest neighbor \\
$\v{B}$ & $\frac{p(p-1)}{2}$ \\ 
\end{tabular}
\end{center}

@@

\section{Gershgorin Circle Theorem}

Reference \url{https://en.wikipedia.org/wiki/Gershgorin_circle_theorem}

Why do we care about this theorem? It provides a sufficient condition to ensure that a square matrix is positive semi-definite.

Let $\v{A}$ be a symmetric $n \times n$ matrix (hence, all eigenvalues are real).
Let $R_{i}=\sum _{j\neq {i}} \left|a_{ij}\right|$ be the sum of the absolute values of the non-diagonal entries in the $i$-th row. Let $D(a_{ii},R_{i})$ be the closed circle (or disc) centered at $a_{ii}$ with radius $R_{i}$. Such a disc is called a Gershgorin disc.

\begin{theorem}[Gershgorin Circle Theorem]
Every eigenvalue of $\v{A}$ lies within at least one of the Gershgorin discs $D(a_{ii},R_{i})$.
\end{theorem}

Let $\lambda^{\min}(\v{A})$ be the smallest eigenvalue of $\v{A}$ and let $\lambda_G^{\min}(\v{A}) = \min_{i=1,\ldots,n} \{a_{ii} - R_i\}$.
\begin{corollary}
If $\lambda_G^{\min}(\v{A}) \geq 0$, then $\v{A}$ is positive semi-definite. 
\end{corollary}
\proof Gershgorin's circle theorem asserts that $\lambda_G^{\min}(\v{A})$ is a lower bound on the minimum eigenvalue of $\v{A}$, i.e., $\lambda^{\min}(\v{A}) \geq \lambda_G^{\min}(\v{A})$.  One definition of a psd matrix is that all of its eigenvalues are non-negative. \qed

In our problem, we may require the matrix $\v{B}$ to be positive semi-definite.  Gershgorin's theorem tells us that this can be partially accomplished using linear programming techniques.  (I say ``partially'' because Gershgorin's theorem is not a necessary condition, thus it enforcing $\lambda_G^{\min}(\v{A}) \geq 0$ is actually a restriction, i.e., it may forbid some eligible psd matrices.  Nevertheless, it's an interesting starting point.)  To accomplish this with our matrix $\v{B}$, let $R_{k}(\v{B})=\sum _{\ell \neq k} \left|b_{k\ell}\right|$ and note that the condition $\lambda_G^{\min}(\v{B}) \geq 0$ implies that $\min_{k=1,\ldots,p} \{b_{kk} - R_k(\v{B})\} \geq 0$, which is equivalent to $b_{kk} - \sum _{\ell \neq k} \left|b_{k\ell}\right| \geq 0$ for all $k=1,\ldots,p$.  After introducing auxiliary non-negative decision variables $b^+_{k\ell}$ to model $|b_{k\ell}|$ for all $k \neq \ell$, the latter can be converted into a set of linear constraint as follows:
\begin{subequations} \label{model:gershgorin_restriction}
\begin{alignat}{4}
& b_{kk} - \sum _{\ell \neq k} b^+_{k\ell} \geq 0 && \qquad \forall k=1,\ldots,p \\
& b^+_{k\ell} \geq b_{k\ell} && \qquad \forall k=1,\ldots,p, \ell=1,\ldots,p~(k \neq \ell) \\
& b^+_{k\ell} \geq -b_{k\ell} && \qquad \forall k=1,\ldots,p, \ell=1,\ldots,p~(k \neq \ell) \\
& b^+_{k\ell} \geq 0 && \qquad \forall k=1,\ldots,p, \ell=1,\ldots,p~(k \neq \ell)
\end{alignat}
\end{subequations}

@@
